Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-04-02 Experiments.tex
5.2 Weak Scalability.
  AA please review/process.

In this section we present experiments % that aim
to determine the performances of the NGE and to show how it provides a minimal
overhead while introducing new functionalities.

Experiments % consist in running
run instances of AthenaMP by using NGE pilots, % where each AthenaMP simulates
simulating a pre-determined number of events in the ATLAS detector.\mtnote{are
they 100 events?}\aanote{`Not always.'}
% taken from ATLAS workload.
We present three groups of experiments in which we test the NGE for weak
scalability, weak scalability with multiple generation, and strong scalability.

We % collected data about
measured the execution time of the pilots and of the AthenaMP % that have been
executed within them, collecting timestamps at % by considering
all stages of the execution. Experiments have been performed % out of production
by % using
submitting NGE's pilots % on
to Titan's batch queue. Because of Titan's % batch queue
policies, the turnaround time of each run of our experiments is dominated by
% the time spent on TITAN's queue
queue time. Since we are interested only in the performances of the NGE, we
removed queue time from our statistics.

All the experiments have been performed by % setting
configuring AthenaMP
% in such a way that it uses
to use all the 16 cores % present on
of Titan's worker nodes.

@@

% The
In this experiment % consists in running
we run as many AthenaMP instances (also referred as tasks from now-on) as the number of nodes controlled by the pilot. Each AthenaMP
% have been assigned with 100 events whose simulation requires
simulates 100 events, requiring $\sim 70$ minutes on average.

Tasks do not % experience
wait within the NGE Agent's queue % within the pilot
since % there is
one node %for
is available to each AthenaMP instance. % Therefore, d
Delays in tasks execution are consequence of % only
three other factors: (i) the bootstrapping of the pilot on the nodes; (ii) the
UnitManager, as defined in Section~\ref{sec:arch}, that has to dispatch tasks to
the agent; and (iii) the time that the agent requires to spawn all the tasks on
the nodes.

We tested % different
pilots % sizes, i.e. :
with 250, 500, 1000 and 2000 worker nodes and
%. For all of them, the walltime was
2 hours walltime. Figure~\ref{fig:weakScal1a} depicts the average pilot
duration, the average execution time of AthenaMP, and the pilot's overhead as
function of the pilot size.

\begin{figure}[!htb]
        \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/weak1.pdf}
    \caption{Weak scalability: average pilot duration, average duration of a
    single AthenaMP execution, and pilot's overhead for different pilot sizes
    (250, 500, 1000 and 2000 worker nodes). }
\label{fig:weakScal1a}
\end{figure}

% Firstly, w
We % can
observe that, despite some fluctuations due to external factors
% \footnote{such as the
(e.g., Titan's shared filesystem and the shared database used by the NGE), the
average execution time of AthenaMP % oscillates
ranges between 4200 and 4800 seconds. % Secondly, w
We % can
also observe that in all the cases the gap between AthenaMP execution times and
the pilot durations is minimal, although it slightly increases with the pilot
size. % By observing NGE's overhead, w
We % can
notice that NGE's overhead does not grow linearly with the number of units.

>>>>

We designed experiments to characterize the performance of the NGE on Titan,
with an emphasis on understanding its overhead and thus the cost of
introducing new functionalities.  We perform three groups of experiments in
which we investigate the weak scalability, weak scalability with multiple
generation, and strong scalability of the NGE.

Each experiment entails executing multiple instances of AthenaMP using the NGE
to simulate a pre-determined number of events. All the experiments have been
performed by  configuring AthenaMP to use all the 16 cores  of Titan's worker
nodes.

We  measured the execution time of the pilots and of the AthenaMP  executed
within them, collecting timestamps at  all stages of the execution.
Experiments have been performed  by  submitting NGE's pilots  to Titan's batch
queue.  The turnaround time of an individual run is determined by queue
waiting times. Since we are interested only in the performances of the NGE, we
removed queue time from our statistics.

@@

In this experiment  we run as many AthenaMP instances (hereafter referred to
as tasks)  as the number of nodes controlled by the pilot. Each AthenaMP
simulates 100 events, requiring $\sim 70$ minutes on average. \jhanote{Please
convert all measurements into seconds!}

Tasks do not  wait within the NGE Agent's queue since  one node  is available
to each AthenaMP instance.  Overheads in task execution are consequence
primarily of the three other factors: (i) the  initial bootstrapping of the
pilot on the nodes; (ii) the UnitManager's dispatching of  units (tasks) to
the agent; and (iii) time for the agent to spawn all the tasks on the nodes.
\jhanote{"spawn" is the right word here. The agent does not spawn tasks. These
are predefined. I would say "initialize" maybe? Something better?}

We tested % different
pilots % sizes, i.e. :
with 250, 500, 1000 and 2000 worker nodes and
%. For all of them, the walltime was
2 hours walltime. The time duration is determined by the Titan's walltime policy. Figure~\ref{fig:weakScal1a} depicts the average pilot
duration, the average execution time of AthenaMP, and the NGE pilot overhead as
function of the pilot size.

\begin{figure}[!htb]
        \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/weak1.pdf}
    \caption{Weak scalability: Average duration of a
    single AthenaMP execution, average pilot duration and pilot's overhead as a function of different pilot sizes. \jhanote{Caption should be in consistent order as the bars in the figure. Ideally pilot first then AthenMP.}}
\label{fig:weakScal1a}
\end{figure}

% Firstly, w
We % can
observe that, despite some fluctuations due to external factors
% \footnote{such as the
(e.g., Titan's shared filesystem and the shared database used by the NGE), the
average execution time of AthenaMP % oscillates
ranges between 4200 and 4800 seconds. % Secondly, w
We % can
also observe that in all the cases the gap between AthenaMP execution times and
the pilot durations is minimal, although it slightly increases with the pilot
size. % By observing NGE's overhead, w
We % can
notice that NGE's overhead does not grow linearly with the number of units.
\jhanote{It does grow linearly with units! What am I missing??}
